# Transparency criteria

__Article:__ _Epidemiological and evolutionary consequences of different types of CRISPR-Cas systems_

__Authors:__ _Hélène Chabas, Viktor Müller, Sebastian Bonhoeffer and Roland R. Regoes_   



The general aim of this study was to determine the importance of molecular differences between 3 types of CRISPR-cas systems. 
The authors developed a stochastic epidemiological model for this.

Outcome would be that type 3 systems could outcompete type 1 and 2. This is because for type 1 and 2 the
control of phage evolution is mediated by spacer diversity and there is a diversity critical threshold whereas for type 3 CRISPR-Cas systems, spacer diversity does not impact phage extinction and the epidemiological outcome is driven by the probability to generate at least one resistant genotype.




```{r transparency criteria}
library(tibble)
score_table <- tibble(`Transparency criteria` = c("Study purpose", "Data availability statement", "Data location", "Study location", "Author review", "Ethics statement", "Funding statement", "Code availability"),
                      Score = c("Yes", "Yes", "Data location not stated", "No", "Tier 4", "No", "Yes", "Yes")
                      
                    
                      )

knitr::kable(score_table)

```




__For trying to reproduce a figure I've decided to use the following article:__
<br>

__Title:__ _Monitoring trends and differences in COVID-19 case fatality rates using decomposition methods: Contributions of age structure and age-specific fatality._

__Authors:__ _Christian Dudel Tim Riffe Enrique Acosta Alyson van Raalte Cosmo Strozza Mikko Myrskylä_



<br><br>
__Question 1:__ Have a look at the code. Describe in your own words what the code intents to achieve.

__Answer 1:__ The code intents to achieve a plot that shows how the death rate progresses as COVID-19 cases rise per country and New York City. 

<br>

__Question 2:__ In terms of readibility of the code, how would you grade (1(very bad)-5(very good)) the code available.

__Answer 2:__ I would grade it a 3. The filtering steps are harder to read for me, but the plotting itself is quite readable. 

<br>

__Question 3:__ Taken together on a scale from 1 (very hard) to 5 (very easy), how much effort did it take you to reproduce the visualization from the project, report or article

__Answer 3:__ I would grade it a 5. The only thing that had to be changed was the path for the inputfile.


```{r plot auteur namaken}
### Monitoring trends and differences in COVID-19 case fatality  ##############
### rates using decomposition methods: A demographic perspective ##############

  ### Last updated: 2020-07-22 11:18:52 CEST
  
  ### Contact:
  ### riffe@demogr.mpg.de
  ### acosta@demogr.mpg.de
  ### dudel@demogr.mpg.de


### Packages ##################################################################

  library(tidyverse)
  library(ggrepel)
  library(scales)
  library(here)


### Load data #################################################################

  # Load data
  db_gh <- read_csv(here("Data/inputdata.csv"))
view(db_gh)

### Aggregate data ############################################################

  # Filter date
  db_gh$Date <- as.Date(db_gh$Date,"%d.%m.%y")
  db_gh2 <- db_gh %>% filter(Date<=as.Date("30.06.2020","%d.%m.%y"))
  
  # Set New York as "country" (easier handling)
  db_gh2$Country[db_gh2$Country=="USA" & db_gh2$Region == "NYC"] <- "NYC"
  
  # Sum data over age groups
  db_gh2 <- db_gh2 %>% 
    filter(!Country %in% c("China","USA","South Korea") & Sex == "b") %>% 
    group_by(Country, Code,Date) %>% 
    summarise(Cases = sum(Cases),
              Deaths = sum(Deaths))

  # Exclude bolletino 
  db_gh2 <- db_gh2 %>%
    filter(str_sub(Code, 1, 5) != "ITbol")
  
  # Sort by date
  db_gh2 <- db_gh2 %>% group_by(Country) %>% arrange(Date)
  
  # Smooth reporting issues cases
  for(country in unique(db_gh2$Country)) {
    
    days <- db_gh2$Date[db_gh2$Country==country]
    
    for(day in 2:length(days)) {
      current <- db_gh2$Cases[db_gh2$Country==country & db_gh2$Date==days[day]]
      previous <- db_gh2$Cases[db_gh2$Country==country & db_gh2$Date==days[day-1]]
      
      if(current<previous) db_gh2$Cases[db_gh2$Country==country & db_gh2$Date==days[day]] <- previous
      
    }
    
  }

  # Smooth reporting issues deaths
  for(country in unique(db_gh2$Country)) {
    
    days <- db_gh2$Date[db_gh2$Country==country]
    
    for(day in 2:length(days)) {
      current <- db_gh2$Deaths[db_gh2$Country==country & db_gh2$Date==days[day]]
      previous <- db_gh2$Deaths[db_gh2$Country==country & db_gh2$Date==days[day-1]]
      
      if(current<previous) db_gh2$Deaths[db_gh2$Country==country & db_gh2$Date==days[day]] <- previous
      
    }
    
  }
  

### Plot settings #############################################################

  # Set colors
  col_country <- c("Germany" = "black",
                   "Italy" = "#2ca25f",
                   "NYC"="#f0027f",
                   "Spain"="#beaed4",
                   "South Korea"="#fdc086")#,
                   #"USA"="#386cb0")
  
  cols <- c("black",
            "#2ca25f",
            "#f0027f",
            "#beaed4",
            "#fdc086")#,
            #"#386cb0")
  
  
  # Axis
  labs <- db_gh2 %>%
    group_by(Country) %>% 
    filter(Cases == max(Cases)) %>% 
    mutate(Cases = Cases + 3000)
  
  # Including all reports
  tx <- 6
  lim_x <- 240000

### Plot ######################################################################

  db_gh2 %>% 
    ggplot(aes(Cases, Deaths, col = Country))+
    geom_line(size = 1, alpha = .9)+
    scale_x_continuous(expand = c(0,0), breaks = seq(0, 300000, 50000), limits = c(0, lim_x + 30000), labels = comma)+
    scale_y_continuous(expand = c(0,0), breaks = seq(0, 40000, 5000), limits = c(0, 40000), labels = comma)+
    annotate("segment", x = 0, y = 0, xend = lim_x, yend = lim_x * .02, colour = "grey40", size = .5, alpha = .3, linetype = 2)+
    annotate("segment", x = 0, y = 0, xend = lim_x, yend = lim_x * .05, colour = "grey40", size = .5, alpha = .3, linetype = 2)+
    annotate("segment", x = 0, y = 0, xend = lim_x, yend = lim_x * .10, colour = "grey40", size = .5, alpha = .3, linetype = 2)+
    annotate("segment", x = 0, y = 0, xend = lim_x, yend = lim_x * .15, colour = "grey40", size = .5, alpha = .3, linetype = 2)+
    annotate("text", label = "2% CFR", x = lim_x + 1000, y = lim_x * .02,
             color="grey30", size = tx * .3, alpha = .6, hjust = 0, lineheight = .8) +
    annotate("text", label = "5% CFR", x = lim_x + 1000, y = lim_x * .05,
             color="grey30", size = tx * .3, alpha = .6, hjust = 0, lineheight = .8) +
    annotate("text", label = "10% CFR", x = lim_x + 1000, y = lim_x * .10,
             color="grey30", size = tx * .3, alpha = .6, hjust = 0, lineheight = .8) +
    annotate("text", label = "15% CFR", x = lim_x + 1000, y = lim_x * .15,
             color="grey30", size = tx * .3, alpha = .6, hjust = 0, lineheight = .8) +
    scale_colour_manual(values = cols)+
    geom_text(data = labs, aes(Cases, Deaths, label = Country),
              size = tx * .35, hjust = 0, fontface = "bold") +
    theme_classic()+
    labs(x = "Cases", 
         y = "Deaths")+
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "none",
      plot.margin = margin(5,5,5,5,"mm"),
      axis.text.x = element_text(size = tx),
      axis.text.y = element_text(size = tx),
      axis.title.x = element_text(size = tx + 1),
      axis.title.y = element_text(size = tx + 1)
    )
  





```
