# Plan on learning beginning steps of scRNA-seq analysis

For the workflows course in the track data science for biology we have been assigned around 32 hours to develop a new skill. In one of the courses we were taught how to analyse data from a bulk RNA-seq experiment. 

Here I am trying to learn how to do the beginning steps of analysing data from a single cell RNA-seq experiment

I have made a planning on how to spend my time trying to learn this new skill, which can be seen here: 

```{r tibble planning, include=F,message=F}
library(tidyverse)
planning <- tibble(Days = c(1, 2,3,4), 
                   Action = c("Reading about single-cell RNA-seq, and finding a dataset that can be used.", "Loading in the data and downloading the required packages + filter data and obtain count matrix.", "Visualising/analyzing the data (DGE)", "See what else can be added, and adjust lay-out of graphs and Rmarkdown accordingly.")
)
```


```{r planning, echo=F}
knitr::kable(planning, caption = "Table 1: My planning for learning to analyse scRNA-seq")
```
Figure 1: Planning for learning scRNA-seq
Days	Action
Day 1	Reading about single-cell RNA-seq, and finding a dataset that can be used.
Day 2	Loading in the data and downloading the required packages + filter data and obtain count matrix.
Day 3	Visualising/analyzing the data (DGE)
Day 4	Adjust lay-out accordingly, and see what else can be added.




useful [source](https://www.singlecellcourse.org/index.html) for learning analysis of scRNA-seq. [github page](https://github.com/cellgeni/scRNA.seq.course)


[useful](https://www.youtube.com/@Seurat_Scanpy_Signac/videos)

[also useful](https://biostatsquid.com/scrnaseq-preprocessing-workflow-seurat/)

[geo dataset](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE22669)


