# Parameterized COVID-19 report

This is a report with three parameters. Parameters are `country`, `year` and `month`. After knitting this Rmarkdown with parameters two plots are generated, one with covid deaths, and one with covid cases in specified country, year and month.

We start with loading the required libraries 

```{r loading libraries, message=F}
library(tidyverse)
library(readr)
library(stringr)
library(tibble)

```

Reading the data from opendata.ecdc.europa.eu, and processing it.

```{r reading datafile and manipulating data}

# Reading data
data <- read.csv("https://opendata.ecdc.europa.eu/covid19/nationalcasedeath_eueea_daily_ei/csv", na.strings = "", fileEncoding = "UTF-8-BOM")

# Tibble for converting monthnumber to monthname 
months <- tibble(nr = c(1:12), name = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))

# Conversion monthnumber to name
month <- months[params$month, "name"] %>% as.character()

# Filtering data for specified country, year and month in parameters.
country <- data %>% filter(countriesAndTerritories == params$country , year == as.integer(params$year) , month == params$month)

# Formatting column dateRep as a date.
country$dateRep <- as.Date(country$dateRep, format = "%d/%m/%Y")

```  


In chunk below two plots are visible. One for COVID-19 cases in  `r params$country` in `r month` of `r params$year`. And one for COVID-19 deaths,

```{r plotting}
country %>% ggplot(aes(x = dateRep, y = cases)) + geom_point() + geom_line() + theme_bw() + 
  labs(title = paste0(params$country, " COVID-19 cases in ",month, " of ", params$year),
       x = "Date",
       y = "Cases")


country %>% ggplot(aes(x = dateRep, y = deaths)) + geom_point() + geom_line() + theme_bw() + labs(title = paste0(params$country, " COVID-19 deaths in ",month, " of ", params$year),
       x = "Date",
       y = "Deaths")

```







